### 2018-02-25 Report/lab-meeting.Rmd
Adds first lab meeting report

---
title: "lab-meeting"
author: "Lukas 'hejtmy' Hejtmanek"
date: "24 February 2018"
output: 
  revealjs::revealjs_presentation:
    theme: night
    transition: none
    self_contained: true
    css: slides.css
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
```

```r

```

# What

---

Transfer of spatial knowledge across differet

# Methods

## Overview

Building has in total XX doors

6 doors are ascribed to fictional workers: Emily, Michael, Daniel, Jesicca, Jacob and Sarah 
 - Taken from the list of most used names in CA from 2000

4 pointing locations

XX Image of their placement

## Design

Full experiemntal design with 4 possible conditions

- vr - vr
- real - vr
- vr - real
- real - real

## Procedure
Participants fill questionnaires

They do the virtualiser training regardless of their condition (even in real-to real)

Experiment has 2 phases, each has two parts: walking/pointing
- Free roaming from door to door - all 6 ofices in 3 blocks (18 trials)
- Pointing from 2 points to all 6 offices; exactly one office is always visible

Drawing blank map

## Settings

6 possible settings

Example of settings

Question: Should we run two same settings? To directly compare performance in the same sequence of goals and points?

# Participants

## Descriptives

## Number crunching
So far we:

- scheduled 56 timeslots
- administered 45 participants
- have `r length(unique(walk_all$id))` usable subjects
- wasted 5 working days

There are some issues with high dropout rate, so far about a 1/3

---

Dropout by gender

---

Dropout by ethnicity

asdasdas

---

Dropout by condition

asdasadas

# Analysis

## Normalisation
Min normalising

# Results

## The bad
Unfortunatelly, out of `r as.numeric(walk_all %>% filter(type == "restimote") %>% count())` trials in the restimote we have `r as.numeric(walk_all %>% filter(type == "restimote") %>% filter(is.na(distance)) %>% count())` not fully executed.

## Distance

Distance absolute
```{r, echo = F,warning=FALSE}
ggplot(walk_all, aes(x = condition, y = distance, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

---

Distance normalised
```{r, echo = F, warning=FALSE}
ggplot(walk_all, aes(x = condition, y = min_norm_distance, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95),geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

---

```{r, echo = T}
walk_all %>% 
  group_by(condition) %>% 
  summarise(p.value = t.test(distance ~ phase)$p.value,
            phase1 = t.test(distance ~ phase)$estimate[[1]],
            phase2 = t.test(distance ~ phase)$estimate[[2]])
```

## Time