### 2018-05-27 Report/arne-conference.Rmd
Adds report statistics

---
title: "Conference output"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "3 May 2018"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
source('../Scripts/analysis-helpers.R')
sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
```

```{r}
walk_all %>% group_by(condition) %>% summarise(N = n_distinct(id))
```

## Walking ----

## Per phase separation
```{r warning=F, echo=F}
make_graph(walk_all, "condition", "distance", "phase") + ylab("Average absolute trial distance") + xlab("condition") + scale_fill_discrete(name = "Phase")
make_graph(walk_all, "condition", "min_norm_distance", "phase") + ylab("Average normalised trial distance") + xlab("condition") + scale_fill_discrete(name = "Phase")
make_graph(walk_all, "condition", "time", "phase") + ylab("Average absolute trial duration") + xlab("condition") + scale_fill_discrete(name = "Phase")
make_graph(walk_all, "condition", "min_norm_time", "phase")+ ylab("Average normalised trial duration") + xlab("condition") + scale_fill_discrete(name = "Phase")
make_graph(walk_all, "condition", "errors", "phase")+ ylab("Average number of errors") + xlab("condition") + scale_fill_discrete(name = "Phase")
```

## PER BLOCK
```{r warning=F, echo=F}
make_graph(walk_all, "condition", "min_norm_distance", "exp_block_id") + ylab("Average normalised trial distance") + xlab("condition") + scale_fill_discrete(name = "Block")
make_graph(walk_all, "condition", "min_norm_time", "exp_block_id")+ ylab("Average normalised trial duration") + xlab("condition") + scale_fill_discrete(name = "Block")
make_graph(walk_all, "condition", "errors", "exp_block_id")+ ylab("Average number of errors") + xlab("condition") + scale_fill_discrete(name = "Block")

ggplot(walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average normalised trial distance") + xlab("Block") + scale_fill_discrete("condition")

ggplot(walk_all, aes(x = exp_block_id, y = errors, color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Number of mistakes in office decision") + xlab("Block") + scale_fill_discrete("condition")

```

## Pointing
```{r warning=F, echo=F}
ggplot(sop_all[sop_all$time < 12, ], aes(x = condition, y = abs(error), fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average pointing error in degrees") + xlab("Phase") + scale_fill_discrete("Condition")

ggplot(sop_all[sop_all$time < 12, ], aes(x = phase, y = abs(error), color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average pointing error in degrees") + xlab("Phase") + scale_fill_discrete("Condition")

ggplot(sop_all[sop_all$time < 12, ], aes(x = condition, y = time, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average pointing time") + xlab("Phase") + scale_fill_discrete("Condition")

ggplot(sop_all[sop_all$time < 12, ], aes(x = phase, y = time, color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average pointing time") + xlab("Phase") + scale_fill_discrete("Condition")

```