### 2018-05-31 Report/arne-conference.Rmd
More arnes comments

---
title: "Conference output"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "3 May 2018"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(reshape2)
library(broom)

source('../Scripts/analysis-helpers.R')
sop_all <- read.table("../sop.csv", sep=";", header = T, stringsAsFactors = F)
walk_all <- read.table("../walk.csv", sep=";", header = T, stringsAsFactors = F)
conference_walk_all <- walk_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
conference_sop_all <- sop_all %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
```

#Number of participants
```{r}
walk_all %>% group_by(condition) %>% summarise(N = n_distinct(id))
```

## Walking ----

## Per phase separation
```{r warning=F, echo=F}
make_graph(conference_walk_all, "condition", "distance", "phase") + ylab("Average absolute trial distance") + xlab("condition") + scale_fill_discrete(name = "Phase")
make_graph(conference_walk_all, "condition", "min_norm_distance", "phase") + ylab("Average normalised trial distance") + xlab("condition") + scale_fill_discrete(name = "Phase")
make_graph(conference_walk_all, "condition", "time", "phase") + ylab("Average absolute trial duration") + xlab("condition") + scale_fill_discrete(name = "Phase")
make_graph(conference_walk_all, "condition", "min_norm_time", "phase")+ ylab("Average normalised trial duration") + xlab("condition") + scale_fill_discrete(name = "Phase")
make_graph(conference_walk_all, "condition", "errors", "phase")+ ylab("Average number of errors") + xlab("condition") + scale_fill_discrete(name = "Phase")
```

## PER BLOCK
```{r warning=F, echo=F}
make_graph(conference_walk_all, "condition", "min_norm_distance", "exp_block_id") + ylab("Average normalised trial distance") + xlab("condition") + scale_fill_discrete(name = "Block")

make_graph(conference_walk_all, "condition", "min_norm_time", "exp_block_id") + ylab("Average normalised trial duration") + xlab("condition") + scale_fill_discrete(name = "Block")

make_graph(conference_walk_all, "condition", "errors", "exp_block_id")+ ylab("Average number of errors") + xlab("condition") + scale_fill_discrete(name = "Block")

ggplot(conference_walk_all, aes(x = exp_block_id, y = min_norm_distance, color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average normalised trial distance") + xlab("Block") + scale_fill_discrete("Condition")

ggplot(conference_walk_all, aes(x = exp_block_id, y = errors, color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Number of mistakes in office decision") + xlab("Block") + scale_fill_discrete("Condition")

```

## Pointing
```{r warning=F, echo=F}
ggplot(conference_sop_all, aes(x = condition, y = abs(error), fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average pointing error in degrees") + xlab("Condition") + scale_fill_discrete("Phase")

ggplot(conference_sop_all, aes(x = phase, y = abs(error), color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average pointing error in degrees") + xlab("Phase") + scale_fill_discrete("Condition")

ggplot(conference_sop_all[conference_sop_all$time < 12, ], aes(x = condition, y = time, fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average pointing time") + xlab("Condition") + scale_fill_discrete("Phase")

ggplot(conference_sop_all[conference_sop_all$time < 12,],aes(x = phase, y = time, color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average pointing time") + xlab("Phase") + scale_fill_discrete("Condition")

```

# Block measures

```{r, echo = F}
condition_all_sums <- conference_walk_all %>% 
  group_by(exp_block_id, id, condition) %>% 
  summarise(sum.distance = sum(min_norm_distance, na.rm = T), 
            sum.errors = sum(errors, na.rm = T),
            mean.distance = mean(min_norm_distance, na.rm = T),
            mean.error = mean(errors, na.rm = T))

block_measure <- function(df, value){
  df_wide <- dcast(df, condition + id ~ exp_block_id, value.var = value)
  colnames(df_wide) <- c("condition","id", "block1","block2","block3","block4","block5","block6")
  df_wide %>% 
    group_by(condition) %>%
    summarise(block1.mean  = mean(block1, na.rm = T),
              block3.mean = mean(block3, na.rm = T),
              block4.mean = mean(block4, na.rm = T),
              block34.diff = mean((block3-block4)/(block4+block3), na.rm = T),
              block34.se  = sqrt(var((block3-block4)/(block4+block3), na.rm = T)/sum(!is.na(block3-block4))),
              block14.diff = mean((block1-block4)/(block1+block4), na.rm = T),
              block14.se = sqrt(var((block1-block4)/(block1+block4), na.rm = T)/sum(!is.na(block1-block4)))
  )
}
```
Block 43 diff is now flipped to (block4 - block3)/(block4 + block3) and block41 is the forth and first.

There were several ways how I could calculate it. Eventually, the block 4 data are sums of errors per person and block and average path is the average normalised path for a block per person.

Measure of overall sum of errors. Mean columns define mean sum of errors for all participants in that block
```{r, echo=F, warnings = F}
df <- block_measure(condition_all_sums, "sum.errors")
df
ggplot(melt(df[,1:4]), aes(x = condition, y = value, fill = variable)) + 
  geom_bar(stat = 'identity', position="dodge") + ylab("Mean normalised distance")

ggplot(df, aes(x = condition, y = block34.diff, fill = condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), 
                width=.2, position=position_dodge(.9)) +
  ylab("Mean sum of errors improvement from block 3 to 4")

ggplot(df, aes(x = condition, y = block14.diff, fill = condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), 
                width=.2, position=position_dodge(.9)) +
  ylab("Mean sum of errors improvement from block 1 to 4")
```

Measure of mean distance. First three columns define mean normalised distance in each block.
```{r, echo=F, warnings = F}
df <- block_measure(condition_all_sums, "mean.distance")
df
ggplot(melt(df[,1:4]), aes(x = condition, y = value, fill = variable)) + geom_bar(stat = 'identity', position="dodge") + ylab("Mean normalised distance")

ggplot(df, aes(x = condition, y = block34.diff, fill = condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), 
                width=.2, position=position_dodge(.9)) +
  ylab("Mean normalised distance improvement from block 3 to 4")

ggplot(df, aes(x = condition, y = block14.diff, fill = condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), 
                width=.2, position=position_dodge(.9)) +
  ylab("Mean normalised distance improvement from block 1 to 4")
```

# Stats

## Block T tests

```{r, echo =F}
block_t_test <- function(df, block1, block2, value){
  df <- df %>% filter(exp_block_id %in% c(block1,block2))
  mean.na <- function(data){return(mean(data, na.rm = T))}
  df_wide <- dcast(df, condition + id ~ exp_block_id, value.var = value, fun.aggregate = mean.na)
  colnames(df_wide) <- c("condition","id", "block1","block2")
  df_wide %>% group_by(condition) %>% do(tidy(t.test(.$block1, .$block2, paired = T)))
}

```

### Block 3-4
Paired t tests between min normalised distance difference in blocks 3 and 4.
```{r}
block_t_test(conference_walk_all, 4, 3,"min_norm_distance")
```

And for number of errors
```{r}
block_t_test(conference_walk_all, 4, 3,"errors")
```

### Block 1 and 4
Paired t tests between min normalised distance difference in blocks 1 and 4.
```{r}
block_t_test(conference_walk_all, 4, 1,"min_norm_distance")
```
And for number of errors
```{r}
block_t_test(conference_walk_all, 4, 1,"errors")
```

### Block 1 and 6
Paired t tests between min normalised distance difference in blocks 1 and 6.
```{r}
block_t_test(conference_walk_all, 6, 1,"min_norm_distance")
```
And for number of errors
```{r}
block_t_test(conference_walk_all, 6, 1,"errors")
```

## Anova
This is simple mixed anova (for those three conditions) I know how to do in R. Not sure how to do contrasts and within add subject error variability, nor how to run post hocs. I expect Mike will be able to do that in his models for the paper.

```{r}
aov_block_distance <- aov(min_norm_distance~exp_block_id*condition, conference_walk_all)
summary(aov_block_distance)
```

## Independent comparisons

```{r, echo=F}
df_wide <- dcast(condition_all_sums, condition + id ~ exp_block_id, value.var = "mean.distance")
colnames(df_wide) <- c("condition","id", "block1","block2","block3","block4","block5","block6")
df_wide$diff14 <- (df_wide$block1-df_wide$block4)/(df_wide$block1+df_wide$block3)
```

1 to 4 improvement comparison between VE-real and real-real
```{r}
t.test(df_wide$diff14[df_wide$condition=="ve-real"],df_wide$diff14[df_wide$condition=="real-real"])

```

1 to 4 improvement comparison between VR-real and real-real
```{r}
t.test(df_wide$diff14[df_wide$condition=="vr-real"],df_wide$diff14[df_wide$condition=="real-real"])  
```

1 to 4 improvement comparison between VR-real and VE-real
```{r}
t.test(df_wide$diff14[df_wide$condition=="vr-real"],df_wide$diff14[df_wide$condition=="ve-real"])
```