### 2019-03-20 Report/first-revision.Rmd
Adds first revision markdown

---
title: "1st Revision"
author: "Lukáš Hejtmánek"
date: "20 March 2019"
output: html_document
---

```{r setup, include=FALSE}
#needs lme and car in the top because of the recode
library(car)
library(lme4)
library(lmerTest)
library(ggplot2)
library(gridExtra)
library(ez)
library(Hmisc)
library(dplyr)
library(reshape2)
library(broom)
library(knitr)
library(papaja)
library(googlesheets)

source("../Scripts/paper-prepare.R")
source("helpers/block-t-test-improvement.R")
source("helpers/block-t-test.R")
source("helpers/block-measure.R")
source("../TwoWorlds/helpers-report.R")

settings <- load_google_sheets()
df_participants <- settings$participants
df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))
all_participant_experiment <- df_participants
df_participants <- df_participants %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))

overview <- settings$versions
overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
all_participant_experiment <- overview %>% left_join(all_participant_experiment[, c('Code', 'sex')], by="Code")
overview <- overview %>% filter(condition %in% c("vr-real", "real-real", "ve-real"))
overview <- overview %>% left_join(df_participants[, c('Code', 'sex')], by="Code")

#green, violet, blue, yellow, pink, dark blue
estimote_pal <- c('#b6d6c1', '#54003d','#93d5f6','#d5d215','#f3aacb','#2d2556')
theme_set(theme_minimal() + theme(axis.text=element_text(size=14),
        axis.title=element_text(size=16)))

### Try running it without outliers
sub_walk_all[!is.na(sub_walk_all$norm_distance) & sub_walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
walk_all[!is.na(walk_all$norm_distance) & walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA

condition_all_sums <- sub_walk_all %>% 
  group_by(exp_block_id, id, learning.condition) %>% 
  summarise(sum.distance = sum(min_norm_distance, na.rm = T), 
            sum.errors = sum(errors, na.rm = T),
            mean.distance = mean(min_norm_distance, na.rm = T),
            mean.error = mean(errors, na.rm = T))

df_errors <- block_measure(condition_all_sums, "sum.errors")
df_distance <- block_measure(condition_all_sums, "mean.distance")
df_distance$graph.learning.condition <- dplyr::recode(df_distance$learning.condition, "Real" = "Real-world", "Treadmill VR" = "Immersive VR")
df_errors$graph.learning.condition <- dplyr::recode(df_errors$learning.condition, "Real" = "Real-world", "Treadmill VR" = "Immersive VR")

df_wide_errors <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "sum.errors")
df_wide_distance <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "mean.distance")
colnames(df_wide_errors) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
colnames(df_wide_distance) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
df_wide_distance$graph.learning.condition <- dplyr::recode(df_wide_distance$learning.condition, "Real" = "Real-world", "Treadmill VR" = "Immersive VR")
df_wide_errors$graph.learning.condition <- dplyr::recode(df_wide_errors$learning.condition, "Real" = "Real-world", "Treadmill VR" = "Immersive VR")

df_wide_errors$block34 <- (df_wide_errors$block3-df_wide_errors$block4)/(df_wide_errors$block3+df_wide_errors$block4)
df_wide_distance$block34 <- (df_wide_distance$block3-df_wide_distance$block4)/(df_wide_distance$block3+df_wide_distance$block4)

sub_walk_all_factor <- sub_walk_all
sub_walk_all_factor$exp_block_id <- factor(sub_walk_all_factor$exp_block_id)
sub_walk_all_factor$id <- factor(sub_walk_all_factor$id)
sub_walk_all_factor$learning.condition <- factor(sub_walk_all_factor$learning.condition)

knitr::opts_chunk$set(echo=F, warning = F, message = F)
```

## Participants
```{r}
# need revising 
# is_ok == data well measured
# finished == did finsih, regardless of potentially missing real world data
finished <- overview %>% group_by(condition) %>% count(finished)
finished <- dcast(finished, condition~finished)
finished$percent <- round(finished$yes/(finished$yes+finished$no), 2)

is_ok <- overview %>% filter(finished=="yes") %>% group_by(condition) %>% count(is_ok)
is_ok <- dcast(is_ok, condition~is_ok)
is_ok$percent <- round(is_ok$yes/(is_ok$yes+is_ok$no), 2)
kable(finished)

finished_gender <- overview %>% group_by(condition, sex) %>% count(finished)
finished_gender <- dcast(finished_gender, condition+sex~finished)
finished_gender$no[is.na(finished_gender$no)] <- 0
finished_gender$percent <- round(finished_gender$yes/(finished_gender$yes+finished_gender$no), 2)
kable(finished_gender)

### All experiment
finished_gender <- all_participant_experiment %>% group_by(condition, sex) %>% count(finished)
finished_gender <- dcast(finished_gender, condition+sex~finished)
finished_gender$no[is.na(finished_gender$no)] <- 0
finished_gender$percent <- round(finished_gender$yes/(finished_gender$yes+finished_gender$no), 2)
kable(finished_gender)

```