### 2019-04-08 Report/first-revision.Rmd

---
title: "1st Revision"
author: "Lukáš Hejtmánek"
date: "20 March 2019"
output: html_document
---

```{r setup, include=FALSE}
#needs lme and car in the top because of the recode
library(car)
library(lme4)
library(ggplot2)
library(gridExtra)
library(effsize)
library(ez)
library(Hmisc)
library(dplyr)
library(reshape2)
library(broom)
library(knitr)
library(papaja)
library(googlesheets)

source("../Scripts/paper-prepare.R")
source("../TwoWorlds/helpers-report.R")
#settings <- load_google_sheets()
load('../settings.data')
df_participants <- settings$participants
df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))

overview <- settings$versions
overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
overview <- overview %>% left_join(df_participants[, c('Code', 'sex')], by="Code")

### Try running it without outliers
walk_all[!is.na(walk_all$norm_distance) & walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
sub_walk_all[!is.na(sub_walk_all$norm_distance) & sub_walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA

knitr::opts_chunk$set(echo=F, warning = F, message = F)
```

## Dropout rate per gender and condition
```{r}
# need revising 
# is_ok == data well measured
# finished == did finsih, regardless of potentially missing real world data
finished <- overview %>% group_by(condition) %>% count(finished)
finished <- dcast(finished, condition~finished)
finished$percent <- round(finished$yes/(finished$yes+finished$no), 2)

is_ok <- overview %>% filter(finished=="yes") %>% group_by(condition) %>% count(is_ok)
is_ok <- dcast(is_ok, condition~is_ok)
is_ok$percent <- round(is_ok$yes/(is_ok$yes+is_ok$no), 2)
kable(is_ok)
kable(finished)

finished_gender <- overview %>% group_by(condition, sex) %>% count(finished)
finished_gender <- dcast(finished_gender, condition+sex~finished)
finished_gender$no[is.na(finished_gender$no)] <- 0
finished_gender$percent <- round(finished_gender$yes/(finished_gender$yes+finished_gender$no), 2)
kable(finished_gender)
```

## POinting results

### Experiment 1 phase 1
```{r}
av <- aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 1, ])
summary(av)
ggplot(sub_sop_all[sub_sop_all$phase == 1, ], aes(learning.condition, abs_error)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

Anova results are `r apa_print(av)$full_result$learning_condition`

### Experiment 1 phase 2
```{r}
av <- aov(abs_error ~ learning.condition, data = sub_sop_all[sub_sop_all$phase == 2, ])
summary(av)
ggplot(sub_sop_all[sub_sop_all$phase == 2, ], aes(learning.condition, abs_error)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```
Anova results are `r apa_print(av)$full_result$learning_condition`

```{r}
ggplot(sub_sop_all, aes(learning.condition, abs_error, group=factor(phase), fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
ggplot(sub_sop_all, aes(factor(phase), abs_error, group=learning.condition, fill=learning.condition)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

### Experiment 2 phase 1
T test results is `r t_test_effect(abs_error ~ learning.condition, data = sub_sop_all_2[sub_sop_all_2$phase == 1, ])`
```{r}
ggplot(sub_sop_all_2[sub_sop_all_2$phase == 1, ], aes(learning.condition, abs_error)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

### Experiment 2 phase 2
T test results is `r t_test_effect(abs_error ~ learning.condition, data = sub_sop_all_2[sub_sop_all_2$phase == 2, ])`
```{r}
ggplot(sub_sop_all_2[sub_sop_all_2$phase == 2, ], aes(learning.condition, abs_error)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

```{r}
ggplot(sub_sop_all_2, aes(learning.condition, abs_error, group=factor(phase), fill=factor(phase))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
ggplot(sub_sop_all_2, aes(factor(phase), abs_error, group=learning.condition, fill=learning.condition)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

### Both experiments split condition
```{r}
av <- aov(abs_error ~ condition, data = sop_all[sop_all$phase == 2, ])
summary(av)
ggplot(sop_all[sop_all$phase == 2, ], aes(condition, abs_error)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```
Anova results are `r apa_print(av)$full_result$condition`

### Both experiments split by final condition
T test results is `r t_test_effect(abs_error ~ type, data = sop_all[sop_all$phase == 2, ])`
```{r}
ggplot(sop_all[sop_all$phase == 2, ], aes(type, abs_error)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

```{r}
ggplot(sop_all, aes(learning.condition, abs_error, group=factor(phase), fill=factor(phase))) +
  facet_wrap(~testing.condition)+
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
ggplot(sop_all, aes(factor(phase), abs_error, group=learning.condition, fill=learning.condition)) +
  facet_wrap(~testing.condition)+
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

### Plot of type to type evolution
```{r}
ggplot(sop_all, aes(phase, abs_error, color=condition)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.25), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.25),geom="line")
```

###
```{r}
ez_point <- sop_all[complete.cases(sop_all),]
ez_point$phase <- as.factor(ez_point$phase)
ezANOVA(ez_point,
        within = .(phase),
        between = .(learning.condition),
        dv=abs_error,
        wid=id
        )
```

## Absolute vs normalised values 
```{r}
b4 <- only_blocks(sub_walk_all, c(4))
apa_print(aov(b4$distance ~ b4$learning.condition))$full_result$b4_learning_condition
apa_print(aov(b4$min_norm_distance ~ b4$learning.condition))$full_result$b4_learning_condition
b4 <- only_blocks(sub_walk_all, c(6))
apa_print(aov(b4$distance ~ b4$learning.condition))$full_result$b4_learning_condition
apa_print(aov(b4$min_norm_distance ~ b4$learning.condition))$full_result$b4_learning_condition
```

## Gender effects
```{r}
walk_all <- walk_all %>% left_join(df_participants[, c('Code', 'sex')], by=c("id"="Code"))
sub_walk_all <- sub_walk_all %>% left_join(df_participants[, c('Code', 'sex')], by=c("id"="Code"))
sop_all <- sop_all %>% left_join(df_participants[, c('Code', 'sex')], by=c("id"="Code"))
sub_sop_all <- sub_sop_all %>% left_join(df_participants[, c('Code', 'sex')], by=c("id"="Code"))
sub_sop_all_2 <- sub_sop_all_2 %>% left_join(df_participants[, c('Code', 'sex')], by=c("id"="Code"))
```

To assess effects of gender, we pooled data per gender and per modality (desktop, immersive VR and real world) and analysed some of the effects separately.

```{r}
ez_data <- ez_prepare(sop_all, 'sex')
ez_fit <- ezANOVA(
        ez_data,
        dv = abs_error,
        wid = id,
        between = .(type, sex),
        return_aov = T,
        type = 3)
```

Analysing complete pointing performance (averaged across both sessions) overall has not been affected by gender `r apa_print(ez_fit$aov)$full_result$sex` or interaction between gender and modality  `r apa_print(ez_fit$aov)$full_result$type_sex`.

### Walkign distance
```{r}
ez_data <- ez_prepare(walk_all, 'sex')
ez_fit <- ezANOVA(
        ez_data,
        dv = min_norm_distance,
        wid = id,
        between = .(type, sex),
        return_aov = T,
        type = 3)
#ez_data %>% group_by(sex, type) %>% summarise(mean = mean(min_norm_distance, na.rm=T))
ggplot(ez_data, aes(graph.type, min_norm_distance, group=sex, fill=sex)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```
Analysing complete walking distance as a product of either gender or modality, (averaged across all blocks) the walked normalised distance has not been affected by gender `r apa_print(ez_fit$aov)$full_result$sex` or interaction between gender and modality  `r apa_print(ez_fit$aov)$full_result$type_sex`.

### Visitation errors
```{r}
ez_data <- ez_prepare(walk_all, 'sex')
ez_fit <- ezANOVA(
        ez_data,
        dv = errors,
        wid = id,
        between = .(type, sex),
        return_aov = T,
        type = 3)
#ez_data %>% group_by(sex, type) %>% summarise(mean = mean(min_norm_distance, na.rm=T))
ggplot(ez_data, aes(graph.type, errors, group=sex, fill=sex)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

Analysing complete visitations errors as a product of either gender or modality, (averaged across all blocks) number of errors has not been affected by gender `r apa_print(ez_fit$aov)$full_result$sex` or interaction between gender and modality  `r apa_print(ez_fit$aov)$full_result$type_sex`.

Therefore we decided not to include gender specific analyses in our paper, as it might muddy the analyses with another predictor.

## New charts

## Effect sizes