### 2019-10-28 Report/second-revision.Rmd
Slight changes, adds second revision results

---
title: "2nd Revision"
author: "Lukáš Hejtmánek"
date: "20 March 2019"
output: html_document
---

```{r setup, include=FALSE}
#needs lme and car in the top because of the recode
library(ggplot2)
library(effsize)
library(ez)
library(Hmisc)
library(dplyr)
library(reshape2)
library(broom)
library(knitr)
library(papaja)
library(googlesheets)

source("../Scripts/paper-prepare.R")
source("../TwoWorlds/helpers-report.R")
load('../settings.data')
df_participants <- settings$participants
df_participants <- df_participants %>% mutate(condition=paste0(First.phase,"-",Second.phase))

overview <- settings$versions
overview <- overview %>% mutate(condition=paste0(First.phase,"-",Second.phase))
overview <- overview %>% left_join(df_participants[, c('Code', 'sex')], by="Code")

### Try running it without outliers
walk_all[!is.na(walk_all$norm_distance) & walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
sub_walk_all[!is.na(sub_walk_all$norm_distance) & sub_walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA

knitr::opts_chunk$set(echo=T, warning = F, message = F)
```

## Difference in experiment 1 block 1

```{r, result="asis"}
apa_print(aov(min_norm_distance ~ learning.condition, data = sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$full_result$learning_condition
apa_print(aov(errors ~ learning.condition, data = sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$full_result$learning_condition
```
## Difference in experiment 1 block 2

```{r}
apa_print(aov(min_norm_distance ~ learning.condition, data = sub_walk_all[sub_walk_all$exp_block_id == 2, ]))$full_result$learning_condition
apa_print(aov(errors ~ learning.condition, data = sub_walk_all[sub_walk_all$exp_block_id == 2, ]))$full_result$learning_condition
```
```{r}
only_blocks(sub_walk_all, 2) %>% group_by(learning.condition) %>% summarize(avg = mean(errors, na.rm=T))
t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), 2))
t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 2))
t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Treadmill VR', 'Desktop'), 2))
```

## Difference in experiment 2 block 1

```{r}
apa_print(aov(min_norm_distance ~ learning.condition, data = sub_walk_all_2[sub_walk_all_2$exp_block_id == 1, ]))$full_result$learning_condition
apa_print(aov(errors ~ learning.condition, data = sub_walk_all_2[sub_walk_all_2$exp_block_id == 1, ]))$full_result$learning_condition
```

## Block 4 comparisons in experiment 1

```{r}
only_blocks(sub_walk_all, 4) %>% group_by(learning.condition) %>% summarize(avg = mean(errors, na.rm=T))
t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), 4))
t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 4))
t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Treadmill VR', 'Desktop'), 4))

only_blocks(sub_walk_all, 4) %>% group_by(learning.condition) %>% summarize(avg = mean(min_norm_distance, na.rm=T))
t_test_effect(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Treadmill VR'), 4))
t_test_effect(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Real', 'Desktop'), 4))
t_test_effect(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all, c('Treadmill VR', 'Desktop'), 4))
```

## Block 4 comparisons in experiment 2

```{r}
only_blocks(sub_walk_all_2, 4) %>% group_by(learning.condition) %>% summarize(avg = mean(errors, na.rm=T))
t_test_effect(errors ~ learning.condition, only_conditions_blocks(sub_walk_all_2, c('Real', 'Treadmill VR'), 4))
only_blocks(sub_walk_all_2, 4) %>% group_by(learning.condition) %>% summarize(avg = mean(min_norm_distance, na.rm=T))
t_test_effect(min_norm_distance ~ learning.condition, only_conditions_blocks(sub_walk_all_2, c('Real', 'Treadmill VR'), 4))
```

## Absolute vs normalised values 
```{r}
b4 <- only_blocks(sub_walk_all, c(4))
apa_print(aov(b4$distance ~ b4$learning.condition))$full_result$b4_learning_condition
apa_print(aov(b4$min_norm_distance ~ b4$learning.condition))$full_result$b4_learning_condition
ggplot(b4, aes(learning.condition, min_norm_distance, fill)) + 
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
b6 <- only_blocks(sub_walk_all, c(6))
apa_print(aov(b6$distance ~ b6$learning.condition))$full_result$b6_learning_condition
apa_print(aov(b6$min_norm_distance ~ b6$learning.condition))$full_result$b6_learning_condition
ggplot(b6, aes(learning.condition, distance, fill)) + 
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar")
```

## Dropout changes
```{r}
finished_condition <- overview %>% filter(Second.phase != "ve") %>% select(First.phase, Second.phase, finished)
finished_condition <- finished_condition %>% mutate(has_vr = First.phase == "vr" |  Second.phase == "vr")
exp_1 <- finished_condition[finished_condition$Second.phase == "vr", ]
chsqt_1 <- table(exp_1$First.phase, exp_1$finished) %>% chisq.test()
exp_2 <- finished_condition[finished_condition$Second.phase == "real", ]
chsqt_2 <- table(exp_2$First.phase, exp_2$finished) %>% chisq.test()
chsqt_1
chsqt_2
```