### 2019-12-12 Report/paper-graphs.Rmd
Adds final results graphs fro the paper modification

---
title: "paper-graphs"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "12/12/2019"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(papaja)

source("../Scripts/paper-prepare.R")
source("../TwoWorlds/helpers-report.R")

#settings <- load_google_sheets()
load('../settings.data')

#green, violet, blue, yellow, pink, dark blue
estimote_pal <- c('#b6d6c1', '#54003d','#93d5f6','#d5d215','#f3aacb','#2d2556')
theme_set(theme_minimal() + theme(axis.text=element_text(size=14),
        axis.title=element_text(size=16)))

### Try running it without outliers
sub_walk_all[!is.na(sub_walk_all$norm_distance) & sub_walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA
walk_all[!is.na(walk_all$norm_distance) & walk_all$norm_distance > 3, c("distance", "min_norm_distance")] <- NA

sub_walk_all_factor <- sub_walk_all
sub_walk_all_factor$exp_block_id <- factor(sub_walk_all_factor$exp_block_id)
sub_walk_all_factor$id <- factor(sub_walk_all_factor$id)
sub_walk_all_factor$learning.condition <- factor(sub_walk_all_factor$learning.condition)

point_data_1 <- sub_sop_all[sub_sop_all$phase == 1,]
knitr::opts_chunk$set(echo=F, warning = F, message = F)
```

## 5a
```{r}
ggplot(only_blocks(sub_walk_all_factor, 1:3), aes(x = graph.exp_block_id, y = errors, fill = graph.exp_block_id)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average visitation errors") + xlab("Learning condition") + 
  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
  theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside")
```

## 5b
```{r}
# Pre transfer plot of normalised distances
ggplot(only_blocks(sub_walk_all_factor, 1:3), aes(x = graph.exp_block_id, y = min_norm_distance, fill = graph.exp_block_id)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average normalised trial distance") + xlab("Learning condition") + 
  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
  theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside")

```

## 5c

```{r}
ggplot(point_data_1, aes(x = graph.learning.condition, y = abs_error, fill = graph.learning.condition)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average pointing error after learning phase") + xlab("Learning condition") + 
  scale_fill_manual(values = estimote_pal, "Learning condition") +
  guides(fill=F) + theme(axis.title.y = element_text(size = rel(0.9)))

```

## 6a
```{r}
ggplot(only_blocks(sub_walk_all_factor, 3:4), aes(x = graph.exp_block_id, y = errors, fill = graph.exp_block_id)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average visitation errors") + xlab("Learning condition") + 
  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
  theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside")
```

## 6b
```{r}
### graphs
ggplot(only_blocks(sub_walk_all_factor, 3:4), aes(x = graph.exp_block_id, y = min_norm_distance, fill = graph.exp_block_id)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average normalised trial distance") + xlab("Learning condition") + 
  scale_fill_manual(values=estimote_pal,"Block") + guides(fill=F) +
  facet_wrap(~graph.learning.condition, strip.position = "bottom", scales = "free_x") + 
  theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside")

```

## 7a

```{r, fig.width=10, fig.height=8}
ggplot(only_blocks(walk_all, 6), aes(x = testing.condition, y = min_norm_distance, fill = graph.learning.condition)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average path distance in the last block of the transfer phase") + xlab("Transfer condition") + 
  scale_fill_manual(values = estimote_pal, "Condition")  + guides(fill=F) + 
  facet_wrap(~graph.learning.condition, strip.position = "top", scales = "free_x") + 
  ggtitle("Learning condition")+
  theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside",
         strip.text = element_text(size = 16),
          plot.title = element_text(hjust = 0.5))

```

##• 7b

```{r}
ggplot(only_blocks(walk_all, 6), aes(x = testing.condition, y = errors, fill = graph.learning.condition)) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(0.95), geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_dodge(width=0.95),geom="bar") +
  ylab("Average number of visitation errors in the last block of the transfer phase") + xlab("Transfer condition") + 
  scale_fill_manual(values = estimote_pal, "Condition")  + guides(fill=F) + 
  facet_wrap(~graph.learning.condition, strip.position = "top", scales = "free_x") + 
  ggtitle("Learning condition")+
  theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside",
         strip.text = element_text(size = 16),
          plot.title = element_text(hjust = 0.5))
```